\documentclass[12pt]{article}

%Packages%
\usepackage{amsmath, bm, xcolor, amssymb, parskip, cancel}

%Meta%
\title{Supervised Learning}
\author{Diego ROJAS}

%Shortcuts%
\newcommand{\R}{{\rm I\!R}}
\newcommand{\N}{{\rm I\!N}}
\newcommand{\bx}{{\bm{x}}}
\newcommand{\A}{{\bm{A}}}
\newcommand{\B}{{\bm{B}}}
\newcommand{\vecs}[2]{{\bm{#1}_1, \dots, \bm{#1}_#2}}
\newcommand{\xdefinition}[2]{\paragraph{\colorbox{#1!30}{\textbf{Definition:}} (#2):}}
\newcommand{\xrule}[1]{\paragraph{\colorbox{#1!30}{\textbf{Rule:}}}}
\newcommand{\mc}[1]{\mathcal{#1}}

%Document%
\begin{document}

\maketitle

%Introduction%	
\section{Introduction}

Supervised learning is a method of machine learning which uses the \textit{inputs} to predict the \textit{outputs}.

\xdefinition{red}{Predictor} Synonym for input.

\xdefinition{red}{Responses $|$ Dependent variables} Synonym for output. Can be a \textit{quantitative} (\textit{discrete}) or \textit{qualitative} (\textit{categorical}) measurement. There also exists \textit{ordered categorical} variables which refer to outputs such as "small", "large", "medium".

\xdefinition{red}{Regression} When we predict quantitative outputs.

\xdefinition{red}{Classification} When we predict qualitative outputs.

\xdefinition{red}{Training data} A dataset $(x_i, y_i)$ or $(x_i, g_i), i = 1, \dots, N$ with which we construct our prediction rules.

\xdefinition{red}{Binary coded target} A mean of representing a 2-valued categorical output $G$ by encoding it as $Y$ and then treat it as a quantitative output.

\section{Linear Model}

\xdefinition{red}{Linear Model} Given a vector of inputs $X^\top = (X_1, \dots, X_p)$ we predict the output $Y$ via the model
%
\begin{equation}
	\hat{Y} = \hat{\beta}_0 + \sum\limits_{j=1}^{p} X_j \hat{\beta}_j
\end{equation}
%
Where $\hat{\beta}_0$ is the \textit{intercept} or \textit{bias}. By including $1$ in $X$ and $\hat{\beta}_0$ in the vector of coefficients $\hat{\beta}$ we can write
%
\begin{equation}
	\hat{Y} = X^\top \hat{\beta}
\end{equation}

\section{Least Squares}

\end{document}